---
title: "preps to run corneto"
output:
  html_document:
    keep_md: yes
    toc: yes
    theme: united
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

# General settings

```{r setup}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  include = TRUE,
  cache = TRUE,
  cache.lazy = FALSE,
  eval = TRUE,
  fig.width = 4 * (1 + sqrt(5)) / 2,
  fig.height = 4,
  dpi = 700
)
knitr::opts_knit$set(root.dir = "/Users/burtsche/Documents/01_repos/kidney_fibrosis//")
```

## Packages

```{r, message=F, warning =F, include=}
library(tidyverse)
library(ggplot2)
library("reshape2")
library(ggrepel)
library(knitr)
library(Biostrings)
library(RColorBrewer)
library(org.Hs.eg.db)
library(OmnipathR)
library(igraph)

mutate <- dplyr::mutate
select <- dplyr::select
group_by <- dplyr::group_by
```

```{r}
options(ggplot2.discrete.colour = c("orange3", "darkslateblue", "darkred", "darkgreen", "darkgrey", "darkviolet"))
options(ggplot2.discrete.fill = c("orange3", "darkslateblue", "darkred", "darkgreen", "darkgrey", "darkviolet"))
```

# COSMOS functions

```{r}
downstream_neighbours <- function(meta_network, n_steps, input_names) {
  meta_g <- graph_from_data_frame(data.frame(s = meta_network$source, t = meta_network$target))

  dn_nbours <- ego(graph = meta_g, order = n_steps, nodes = input_names, mode = "out")

  sub_nodes <- c(unique(names(unlist(dn_nbours))), input_names)

  meta_network <- meta_network[meta_network$source %in% sub_nodes & meta_network$target %in% sub_nodes, ]

  return(meta_network)
}


downstream_neighbours2 <- function(meta_network, n_steps, input_names) {
  meta_g <- igraph::graph_from_data_frame(data.frame(s = meta_network$source, t = meta_network$target, i = meta_network$mor), directed = TRUE)

  up_nbours <- igraph::ego(graph = meta_g, order = n_steps, nodes = input_names, mode = "in")

  sub_nodes <- c(unique(names(unlist(up_nbours))), input_names)

  meta_network <- meta_network[meta_network$source %in% sub_nodes & meta_network$target %in% sub_nodes, ]

  return(meta_network)
}



filter_TF_sign <- function(network, gene_expression_binarized, signaling_data, tf_regulon) {
  gene_exp_df <- data.frame(
    gene = gene_expression_binarized$gene,
    target_sign = gene_expression_binarized$signFC
  )

  signaling_df <- data.frame(
    TF = signaling_data$source,
    TF_sign = sign(signaling_data$score)
  ) %>%
    dplyr::filter(.data$TF %in% tf_regulon$tf)

  annotated_network <- network %>%
    dplyr::left_join(
      gene_exp_df,
      by = c(target = "gene")
    ) %>%
    dplyr::left_join(
      signaling_df,
      by = c(source = "TF")
    ) %>%
    dplyr::mutate(source_is_TF = .data$source %in% tf_regulon$tf)

  annotated_network <- annotated_network %>%
    dplyr::mutate(target_gene_unchanged = .data$target_sign == 0) %>%
    dplyr::mutate(TF_target_unchanged = .data$source_is_TF & (.data$target_gene_unchanged | is.na(.data$target_gene_unchanged))) %>%
    dplyr::mutate(inconsistent_TF_gene_sign = sign(.data$TF_sign) != interaction * sign(.data$target_sign))

  removed_interactions <- annotated_network %>%
    dplyr::filter(.data$TF_target_unchanged | .data$inconsistent_TF_gene_sign)

  kept_interactions <- annotated_network %>%
    dplyr::filter(!.data$TF_target_unchanged | is.na(.data$TF_target_unchanged)) %>%
    dplyr::filter(!.data$inconsistent_TF_gene_sign | is.na(.data$inconsistent_TF_gene_sign))

  out_pkn <- kept_interactions %>%
    dplyr::select(
      .data$source,
      .data$interaction, .data$target
    )

  return(out_pkn)
}
```


#1. Load data

```{r}
load("results/2024-07-24_tf_enrichment_results.RData")
load("results/2024-06-12_kinase_enrichment_result.RData")
load("data/processed_data/2024-07-24_diff_results.RData")
```


## Secreted proteins

Get a list of all proteins actually secreted

```{r}
secreted_proteins <- msigdbr::msigdbr(species = "Homo sapiens") %>%
  filter(gs_id %in% c("M5889", "M5885"))
```

### additional proteomics hits

```{r}
additional_secretome <- bind_rows(
  data.frame(
    id = c("COL1A1", "COL5A1", "SERPINE2", "SPARC", "ITGB1", "VIM", "JUP", "ACTA1", "HSPG2", "LOXL2", "TNC", "TGFBI", "IGFBP3", "IGFBP7", "LTBP2", "TAGLN", "CCN2", "LRATD2", "MRC2", "FN1", "FBN1", "BGN", "ADGRG1", "MMP2", "ITGA11", "AMIGO2", "ADAM12"),
    score = 1
  ),
  data.frame(
    id = c("CPA4", "DCDC2", "PLEKHG4", "ITGB1BP1"),
    score = -1
  )
)
```

### fix TGFb signaling module

```{r}
additional_edges <- data.frame(
  source = "TGFB1",
  mor = 1,
  target = c("AKT1", "PI3K", "MAPK1", "SMAD1", "SMAD2", "SMAD3", "SMAD4", "SMAD5", "MAPK14")
)
```

#2. Network 1

##A Get PKN

```{r}
pkn_df <- as.data.frame(import_all_interactions()) %>%
  # remove corrupted resource
  filter(!(sources == "Wang") &
    # use only interaction with clear consensus signal
    (abs(consensus_stimulation) > 0 | abs(consensus_inhibition > 0))) %>%
  # determine clean sign
  mutate(sign = consensus_stimulation - consensus_inhibition) %>%
  # bring in right format
  select(source = source_genesymbol, mor = sign, target = target_genesymbol) %>%
  distinct(source, mor, target) %>%
  # filter for expressed interactions only
  filter(target %in% unique(diff_results$feature_id) & source %in% unique(diff_results$feature_id)) %>%
  filter(mor != 0)
```

##B Filter enzymes and secretome

```{r}
enzymes_df <-
  res_kinase_enrichment$enrichment %>%
  filter(p_value < 0.03 & abs(score) > 3 & time %in% c("0.08h", "1h", "12h")) %>%
  mutate(abs_score = abs(score)) %>%
  group_by(source) %>%
  mutate(max_score = max(abs(score))) %>%
  filter(max_score == abs_score) %>%
  distinct(source, score) %>%
  bind_rows(res_tf_enrichment$enrichment %>%
    filter(p_value < 0.03 & abs(score) > 3 & time %in% c("0.08h", "1h", "12h")) %>%
    mutate(abs_score = abs(score)) %>%
    group_by(source) %>%
    mutate(max_score = max(abs(score))) %>%
    filter(max_score == abs_score) %>%
    distinct(source, score)) %>%
  ungroup() %>%
  drop_na() %>%
  filter(source %in% c(pkn_df$source, pkn_df$target))

secretome_df <- diff_results %>%
  # get early secretomics hits
  # make sure they are atctually secreted proteins
  filter(modality == "secretomics" &
    adj.P.Val < 0.05 &
    abs(logFC) > log2(1.5) &
    time %in% c("0.08h", "1h", "12h", "24h") &
    feature_id %in% secreted_proteins$gene_symbol | feature_id %in% additional_secretome$id) %>%
  # calculate consensus score from all included early timepoints
  mutate(abs_score = abs(logFC), id = feature_id) %>%
  group_by(id) %>%
  mutate(max_score = max(abs(logFC))) %>%
  filter(max_score == abs_score) %>%
  distinct(id, score = logFC) %>%
  ungroup() %>%
  drop_na() %>%
  # remove TGF as it is the model starting point
  filter(id != "TGFB1") %>%
  # remove any enzymes that are upstream in the model (none last time I checked)
  filter(!(id %in% enzymes_df$source)) %>%
  distinct(id, .keep_all = T) %>%
  filter(id %in% c(pkn_df$source, pkn_df$target))
```

##C PKN parsing

###Downstream neighbours

```{r}
pkn_df1 <- downstream_neighbours(pkn_df %>% select(source, target, mor), 5, unique(c(enzymes_df$source, secretome_df$id)))

enzymes_df <- enzymes_df %>% filter(source %in% c(pkn_df1$source, pkn_df1$target))
secretome_df <- secretome_df %>% filter(id %in% c(pkn_df1$source, pkn_df1$target))

pkn_df2 <- downstream_neighbours2(pkn_df1 %>% select(source, target, mor), 5, unique(c(enzymes_df$source, secretome_df$id)))

enzymes_df <- enzymes_df %>% filter(source %in% c(pkn_df2$source, pkn_df2$target))
secretome_df <- secretome_df %>% filter(id %in% c(pkn_df2$source, pkn_df2$target))
```

### TF sign filtering

```{r}
expression_df <- diff_results %>%
  filter(modality == "rna" | modality == "proteomics" | modality == "secretomics") %>%
  filter(time %in% c("0.08h", "1h", "12h", "24h")) %>%
  group_by(modality, feature_id) %>%
  summarise(signFC = sign(mean(logFC))) %>%
  group_by(feature_id) %>%
  summarise(signFC = sign(sum(signFC, na.rm = T))) %>%
  select(gene = feature_id, signFC)

pkn_df3 <- filter_TF_sign(
  network = pkn_df2 %>% dplyr::rename(interaction = mor),
  gene_expression_binarized = expression_df,
  signaling_data = enzymes_df,
  tf_regulon = res_tf_enrichment$net_dorothea %>% select(tf = source, target = target, sign = mor)
) %>%
  dplyr::rename(mor = interaction) %>%
  # fix TGF pathway module
  filter(!grepl("SMAD", target)) %>%
  bind_rows(additional_edges) %>%
  distinct(source, mor, target)

enzymes_df <- enzymes_df %>% filter(source %in% c(pkn_df3$source, pkn_df3$target))
secretome_df <- secretome_df %>% filter(id %in% c(pkn_df3$source, pkn_df3$target))
```



#4. Save for optimisation

Input needed for corneto

- pkn_df.tsv comprised of source, mor, target
- enzymes_df.tsv comprised of source and score

```{r}
pkn_df3 %>% write_tsv("network_data/2024-06-27/pkn_df.tsv")

# %>%filter(!grepl("SMAD", target)) %>%
#   bind_rows(additional_edges) %>%
#   distinct(source, mor, target)


enzymes_df %>% write_csv("network_data/2024-06-27/enzymes_df.csv")
secretome_df %>% write_csv("network_data/2024-06-27/secretome_df.csv")
```

### for visualisation

```{r}
res_network_inputs <- list()

res_network_inputs$enzymes_df_early <- enzymes_df
res_network_inputs$secretome_df_early <- secretome_df
res_network_inputs$pkn_df_early <- pkn_df2

# this is saved at the end with the late inputs
```

#3. Network 2

## A Extract info from first run

```{r}
result_first_df <- read_tsv("network_data/2024-06-27/combinededges_df.tsv") %>%
  separate(edge, into = c("Node1", "sign", "Node2"), remove = F, sep = "--")

early_secretome <- secretome_df %>%
  filter(id %in% unique(c(result_first_df$Node1, result_first_df$Node2)))
```

## B Get PKN

```{r}
pkn_df <- as.data.frame(import_all_interactions()) %>%
  # remove corrupted resource
  filter(!(sources == "Wang") &
    # use only interaction with clear consensus signal
    (abs(consensus_stimulation) > 0 | abs(consensus_inhibition > 0))) %>%
  # determine clean sign
  mutate(sign = consensus_stimulation - consensus_inhibition) %>%
  # bring in right format
  select(source = source_genesymbol, mor = sign, target = target_genesymbol) %>%
  distinct(source, mor, target) %>%
  # filter for expressed interactions only
  filter(target %in% unique(diff_results$feature_id) & source %in% unique(diff_results$feature_id)) %>%
  filter(mor != 0)
```

## C Filter enzymes and secretome

```{r}
perturbation_df <- early_secretome %>%
  bind_rows(data.frame(id = "TGFB1", score = 1))

enzymes_df <-
  res_kinase_enrichment$enrichment %>%
  filter(p_value < 0.03 & abs(score) > 3 & time %in% c("24h", "48h", "72h", "96h")) %>%
  mutate(abs_score = abs(score)) %>%
  group_by(source) %>%
  mutate(max_score = max(abs(score))) %>%
  filter(max_score == abs_score) %>%
  distinct(source, score) %>%
  bind_rows(res_tf_enrichment$enrichment %>%
    filter(p_value < 0.03 & abs(score) > 3 & time %in% c("24h", "48h", "72h", "96h")) %>%
    mutate(abs_score = abs(score)) %>%
    group_by(source) %>%
    mutate(max_score = max(abs(score))) %>%
    filter(max_score == abs_score) %>%
    distinct(source, score)) %>%
  ungroup() %>%
  drop_na() %>%
  filter(source %in% c(pkn_df$source, pkn_df$target))

secretome_df <- diff_results %>%
  # get early secretomics hits
  # make sure they are atctually secreted proteins
  filter(modality == "secretomics" &
    adj.P.Val < 0.05 &
    abs(logFC) > log2(1.5) &
    time %in% c("48h", "72h", "96h") &
    feature_id %in% secreted_proteins$gene_symbol | feature_id %in% additional_secretome$id) %>%
  # calculate consensus score from all included early timepoints
  mutate(abs_score = abs(logFC), id = feature_id) %>%
  group_by(id) %>%
  mutate(max_score = max(abs(logFC))) %>%
  filter(max_score == abs_score) %>%
  distinct(id, score = logFC) %>%
  ungroup() %>%
  drop_na() %>%
  # remove TGF as it is the model starting point
  filter(id != "TGFB1") %>%
  # remove any enzymes that are upstream in the model (none last time I checked)
  filter(!(id %in% enzymes_df$source)) %>%
  distinct(id, .keep_all = T) %>%
  filter(id %in% c(pkn_df$source, pkn_df$target))
```

##D PKN parsing

### Downstream neighbours

```{r}
pkn_df1 <- downstream_neighbours(pkn_df %>% select(source, target, mor), 5, unique(c(enzymes_df$source, secretome_df$id, perturbation_df$id)))

enzymes_df <- enzymes_df %>% filter(source %in% c(pkn_df1$source, pkn_df1$target))
secretome_df <- secretome_df %>% filter(id %in% c(pkn_df1$source, pkn_df1$target))
perturbation_df <- perturbation_df %>% filter(id %in% c(pkn_df1$source, pkn_df1$target))

pkn_df2 <- downstream_neighbours2(pkn_df1 %>% select(source, target, mor), 5, unique(c(enzymes_df$source, secretome_df$id, perturbation_df$id)))

enzymes_df <- enzymes_df %>% filter(source %in% c(pkn_df2$source, pkn_df2$target))
secretome_df <- secretome_df %>% filter(id %in% c(pkn_df2$source, pkn_df2$target))
perturbation_df <- perturbation_df %>% filter(id %in% c(pkn_df2$source, pkn_df2$target))
```

### TF sign filtering

```{r}
expression_df <- diff_results %>%
  filter(modality == "rna" | modality == "proteomics" | modality == "secretomics") %>%
  filter(time %in% c("72h", "96h")) %>%
  group_by(modality, feature_id) %>%
  summarise(signFC = sign(mean(logFC))) %>%
  group_by(feature_id) %>%
  summarise(signFC = sign(sum(signFC, na.rm = T))) %>%
  select(gene = feature_id, signFC)

pkn_df3 <- filter_TF_sign(
  network = pkn_df2 %>% dplyr::rename(interaction = mor),
  gene_expression_binarized = expression_df,
  signaling_data = enzymes_df,
  tf_regulon = res_tf_enrichment$net_dorothea %>% select(tf = source, target = target, sign = 1)
) %>%
  dplyr::rename(mor = interaction) %>%
  # fix TGF pathway module
  filter(!grepl("SMAD", target)) %>%
  bind_rows(additional_edges) %>%
  distinct(source, mor, target)

enzymes_df <- enzymes_df %>% filter(source %in% c(pkn_df3$source, pkn_df3$target))
secretome_df <- secretome_df %>% filter(id %in% c(pkn_df3$source, pkn_df3$target))
perturbation_df <- perturbation_df %>% filter(id %in% c(pkn_df3$source, pkn_df3$target))
```

##E Save for optimisation

Input needed for corneto

- pkn_df.tsv comprised of source, mor, target
- enzymes_df.tsv comprised of source and score
- secretomics df
- perturbation df

```{r}
pkn_df3 %>% write_tsv("network_data/2024-06-27/pkn_df_late.tsv")

# %>%filter(!grepl("SMAD", target)) %>%
#   bind_rows(additional_edges) %>%
#   distinct(source, mor, target)


enzymes_df %>% write_csv("network_data/2024-06-27/enzymes_df_late.csv")
secretome_df %>% write_csv("network_data/2024-06-27/secretome_df_late.csv")
perturbation_df %>% write_csv("network_data/2024-06-27/perturbation_df_late.csv")
```

### for visualisation

```{r}
res_network_inputs$enzymes_df_late <- enzymes_df
res_network_inputs$secretome_df_late <- secretome_df
res_network_inputs$pkn_df_late <- pkn_df2
```

```{r}
save(res_network_inputs, file = "network_data/2024-06-27/networkinput_object.RData")
```
