---
title: "Image analysis"
output:
  html_document:
    keep_md: yes
    toc: yes
    theme: united
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

# General settings

```{r setup}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  include = TRUE,
  cache = TRUE,
  cache.lazy = FALSE,
  eval = TRUE,
  fig.width = 4 * (1 + sqrt(5)) / 2, 
  fig.height = 4,
  dpi = 700
)
```

## Packages

```{r, message=F, warning =F, include=}
library(tidyverse)
library(ggplot2);theme_set(cowplot::theme_cowplot(font_size = 15) + theme(panel.grid.major = element_line(colour = "lightgrey", linewidth = 0.2), panel.grid.minor = element_line(colour = "lightgrey", linewidth = 0.2)))
library("reshape2")
library(ggrepel)
library(knitr)
library(Biostrings)
library(RColorBrewer)
library(org.Hs.eg.db)
library(lme4)

mutate <- dplyr::mutate
select <- dplyr::select
group_by <- dplyr::group_by

tilted <-  theme(axis.text.x = element_text(angle = 45, hjust =1, vjust = 1))
```

# Initial ECM monitoring

```{r}
data <- read.csv(("data/NT22_004_Col1_20240127.csv"))
```

## LMM - discrete time
- calculate mean of all images per sample
- duplicate 0h ctrl for TGFb. There is no 0h treatment but we need it for the model.
- square root transform data to stabilize the variance (decided based on residuals plots) (Piepho 2009)
- remove unstained condition
```{r}
model_dat_n <- data %>%
  mutate(wellID = paste(plate,well, sep = "_")) %>%
  group_by(plate, time, condition, replicates) %>% 
  summarize(mean_int = mean(Col1_per_cell)) %>% # calculate mean per sample (among 36 images)
  mutate(sqrt_mean_int = sqrt(mean_int)) %>% # sqrt transform
  #mutate(log_mean_int = log(mean_int)) %>%
  ungroup() %>%
  filter(condition != "unstained") # this was only needed for autofluorescence correction

# Filter rows with time == 0h
filtered_rows <- model_dat_n %>% filter(time == "0h")

# Duplicate rows
duplicated_rows <- filtered_rows %>% dplyr::slice(rep(1:n(), each = 1))

# Rename condition to "TGF" in duplicated rows
duplicated_rows$condition <- "TGF"

# Combine the original and duplicated rows
model_dat_discreteT_mean <- rbind(model_dat_n, duplicated_rows)
```

```{r linear mixed model discrete time}
lme_model_discrete_mean <- lme4::lmer(formula = sqrt_mean_int ~ condition*time + (1|plate) , data = model_dat_discreteT_mean)

summary(lme_model_discrete_mean)
confint(lme_model_discrete_mean) # confidence intervals
ranef(lme_model_discrete_mean) # to call random effects estimates
```

```{r}

df <- lme4::fortify.merMod(lme_model_discrete_mean, data= lme4::getData(lme_model_discrete_mean)) %>% 
  mutate(time = as.numeric(str_replace(time, "h", "")))

df %>% 
 ggplot(aes(time, sqrt_mean_int, colour = condition, fill = condition))+
  #geom_boxplot() +
  geom_point(aes(shape = plate), size = 3, alpha = 0.8) +
  geom_smooth(method = "lm",se = T, width = 0.1) +
  #ggpubr::stat_compare_means(method = "wilcox.test", label = "p.signif") +
  scale_color_manual(values = c("grey90", "#FED395FF")) +
  scale_fill_manual(values = c("grey90", "#FED395FF")) +
  labs(y = "COLI - norm fluorescence intensity")
```



```{r}
newdata <- data.frame(time= rep("0h",4),
                      condition = rep("ctrl",4),
                      plate = c("A","B","C","D"))

norm_data <- newdata %>% 
  mutate(norm_factor = predict(lme_model_discrete_mean, newdata=newdata)) %>% 
  dplyr::select(plate, norm_factor)


# plot
data_norm_model <- model_dat_discreteT_mean %>% 
  mutate(pred = predict(lme_model_discrete_mean)) %>% 
  left_join(norm_data) %>% 
  mutate(norm_sqrt_mean_int = sqrt_mean_int - norm_factor) %>% 
  mutate(pred_norm = pred - norm_factor) %>% 
  mutate(grouping = paste(plate,condition,sep="")) %>% 
   mutate(time = as.numeric(str_replace(time, "h", "")))

data_norm_model %>%
  ggplot(aes(x=time, y= norm_sqrt_mean_int, col=condition, fill = condition, group = condition)) + 
  geom_jitter(size = 3, alpha = 0.8, width = 0.8) +
  geom_smooth(method = "lm",se = T, width = 0.1) +
  #ggpubr::stat_compare_means(method = "wilcox.test", label = "p.signif") +
  scale_color_manual(values = c("grey90", "#FEA873FF")) +
  scale_fill_manual(values = c("grey90", "#FEA873FF")) +
  labs(y = "COLI - norm fluorescence intensity")
```


```{r}
imaging_data <- list(
  "plot_data" = data_norm_model,
  "model_data" = lme_model_discrete_mean
)
save(imaging_data, file = paste0("results/", Sys.Date(), "_imaging_data_fig1.RData"))
```

# Validations

```{r}
df_outliersEXnew <- read_csv("data/NT24_001_CNA35_20240224.csv")
```
ratio of TGF vs ctrl 
ratio of ratio is the ratio of e.g. (siTranscriptionFactor TGF vs siTranscriptionFactor ctrl) vs (siNeg9 TGF vs siNeg9 ctrl)

create df with intensity_norm_nuclei number values for each control
```{r}
control_df <- df_outliersEXnew %>%
  filter(siRNA == "siNeg9" & condition == "ctrl") %>%
  group_by(plate) %>%
  summarise(ctrl_mean_intensity = mean(Col1_per_cell))

df_test_TGF <- df_outliersEXnew %>%
  #filter(condition == "TGF") %>%
  left_join(control_df, by = c("plate")) %>%
  mutate(FC_intensity = (Col1_per_cell - ctrl_mean_intensity) / ctrl_mean_intensity)

```

## ratio TGF
```{r}
# Calculate intercept for condition "TGF"
intercept_TGF <- df_test_TGF %>%
  filter(#plate %in% c("A","B","C"), 
         siRNA == "siNeg9", condition == "TGF") %>%
  summarise(median_TGF = median(FC_intensity)) %>%
  pull(median_TGF)

# Calculate intercept for condition "ctrl"
intercept_ctrl <- df_test_TGF %>%
  filter(#plate %in% c("A","B","C"),
         siRNA == "siNeg9", condition == "ctrl") %>%
  summarise(median_ctrl = median(FC_intensity)) %>%
  pull(median_ctrl)
```

```{r}
library(ggforce)
df_test_TGF %>%
  filter(
    siRNA %in% c("siNeg9","siE2F1a", "siE2F1b", "siFLI1a", "siFLI1b","siSMAD1a", "siSMAD1b", "siNR4A1", "siHNF4G",  "siBHLHE40")
    ) %>%
  filter(!(condition == "ctrl" & siRNA != "siNeg9")) %>% 
  mutate(siRNA = ifelse(siRNA == "siNeg9", paste(siRNA, condition), siRNA)) %>% 
  mutate(siRNA = factor(siRNA, levels = c("siNeg9 ctrl", "siNeg9 TGF","siE2F1a", "siE2F1b", "siFLI1a", "siFLI1b","siSMAD1a", "siSMAD1b", "siNR4A1", "siHNF4G",  "siBHLHE40"))) %>% 
  ggplot(aes(x = siRNA, y = FC_intensity)) +
  geom_hline(yintercept = intercept_TGF, color = "gray26", linetype = 4, size = 1) +
  geom_hline(yintercept = intercept_ctrl, color = "gray26", size = 1) +
  geom_violin(fill = "white", colour = "mediumvioletred") +
  geom_sina(aes(fill = siRNA), size = 3, alpha = 0.6, shape = 21, colour = "grey")+
  stat_summary(geom = "crossbar", fun= median, color = "black", linewidth = 0.4)+
  ggpubr::stat_compare_means(ref.group = "siNeg9 TGF", label = "..p.signif..", method = 't.test', hide.ns = T, size = 6, label.y = 7.7)+
  scale_fill_manual(values = c("siNeg9" = "darkgrey",
                                "siNR4A1" = "mediumvioletred", #424242
                                "siHNF4G" = "mediumvioletred", #757575
                                "siBHLHE40" ="mediumvioletred", #9E9E9E
                                "siE2F1a" = "mediumvioletred",
                                "siE2F1b" = "mediumvioletred",
                                "siFLI1a" = "mediumvioletred",
                                "siFLI1b" = "mediumvioletred",
                                "siSMAD1a"= "mediumvioletred",
                                "siSMAD1b" = "mediumvioletred")) +
  #facet_wrap(~ condition)+
  ylab("Col1/cell norm to siNeg9 ctrl")+
  ylim(-1,8) +
  #cowplot::panel_border()+
  tilted
```


```{r}
df_test_TGF %>%
  mutate(siRNA = factor(siRNA, levels = c(
   "siNeg9",  "siFN1", "siSMAD1a", "siSMAD1b", "siSMAD1c", "siNR4A1", "siHNF4G", "siBHLHE40", "siE2F1a", "siE2F1b", "siFLI1a", "siFLI1b", "siCOL1A175", "siCOL1A176", "siCOL1A177"))) %>% 
  ggplot(aes(x = siRNA, y = FC_intensity)) +
  geom_hline(yintercept = intercept_TGF, color = "gray26", linetype = 4, size = 1) +
  geom_hline(yintercept = intercept_ctrl, color = "gray26", size = 1) +
  ggforce::geom_sina(fill = "grey", size = 3, alpha = 0.6, shape = 21, colour = "grey")+
  stat_summary(geom = "crossbar", fun= median, color = "black", linewidth = 0.4)+
  ggpubr::stat_compare_means(ref.group = "siNeg9", label = "..p.signif..", method = 't.test', hide.ns = T, size = 6, label.y = 7.7)+
  facet_wrap(~ condition)+
  ylab("Col1/cell norm to siNeg9 ctrl")+
  ylim(-1,8) +
  cowplot::panel_border()+
  tilted
```


# Save

```{r}
res_imaging_validations <- list("data" = df_test_TGF, "intercepts" = list("TGF" = intercept_TGF,"ctl" = intercept_ctrl))
save(res_imaging_validations, file = paste0("results/", Sys.Date(), "_imaging_validations_finaldf.RData"))
```







